---
title: "Incremental validity - A caution"
bibliography: references.bib
biblio-style: apalike
link-citations: yes
params: 
    SHOW_SOLS: TRUE
    TOGGLE: TRUE
---

```{r setup, include=FALSE}
source('assets/setup.R')
library(tidyverse)
library(patchwork)
```

# Prelude

A common goal for researchers is to determine which variables matter (and which do not) in contributing to some outcome variable. A common approach to answer such questions is to consider whether some variable $X$'s contribution remains significant _after_ controlling for variables $Z$.  

The reasoning:  

+ If our measure of $X$ correlates significantly with outcome $Y$ even when controlling for our measure of $Z$, then $X$ contributes to $y$ *over and above* the contribution of $Z$. 
    
  
In multiple regression, we might fit the model $Y = b_0 + b_1 \cdot X + b_2 \cdot Z + \epsilon$ and conclude that $X$ is a useful predictor of $Y$ *over and above* $Z$ based on the estimate $\hat b_1$, or via model comparison between that model and the model without $Z$ as a predictor ($Y = b_0 + b_1 \cdot X + \epsilon$).  
  
# Toy Example
  
Suppose we have monthly data over a seven year period which captures the number of shark attacks on swimmers each month, and the number of ice-creams sold by beach vendors each month.  
Consider the relationship between the two:  
```{r echo=FALSE}
read_csv("https://uoepsy.github.io/data/sharks.csv") %>%
ggplot(.,aes(x=ice_cream_sales,y=shark_attacks))+geom_point()+stat_smooth(method="lm")
```

We can fit the linear model and see a significant relationship between ice cream sales and shark attacks:  
```{r}
sharkdata <- read_csv("https://uoepsy.github.io/data/sharks.csv")
shark_mdl <- lm(shark_attacks ~ ice_cream_sales, data = sharkdata)
summary(shark_mdl)
```

`r qbegin()`
Does the relationship between ice cream sales and shark attacks make sense? What might be missing from our model? 
`r qend()`
`r solbegin(show=TRUE, toggle=FALSE)`
You might quite rightly suggest that this relationship is actually being driven by temperature - when it is hotter, there are more ice cream sales _and_ there are more people swimming (hence more shark attacks). 
`r solend()`


`r qbegin()`
Is $X$ (the number of ice-cream sales) a useful predictor of $Y$ (numbers of shark attacks) over and above $Z$ (temperature)?  
<br>
We might answer this with a multiple regression model including both temperature and ice cream sales as predictors of shark attacks: 
```{r}
shark_mdl2 <- lm(shark_attacks ~temperature + ice_cream_sales, data = sharkdata)
anova(shark_mdl2)
```
<br>
What do you conclude?

`r qend()`
`r solbegin(show=TRUE, toggle=FALSE)`
It appears that numbers of ice cream sales is _not_ a significant predictor of sharks attack numbers over and above the temperature.  
`r solend()`

__However...__
In psychology, we can rarely observe and directly measure the constructs which we are interested in (for example, personality traits, intelligence, emotional states etc.). We rely instead on measurements of, e.g. behavioural tendencies, as a proxy for personality traits.  

Let's suppose that instead of including temperature in degrees celsius, we asked a set of people to self-report on a scale of 1 to 7 how hot it was that day. This measure should hopefully correlate well with the _actual_ temperature, however, there will likely be some variation: 
```{r echo=FALSE}
ggplot(data = sharkdata, aes(x = temperature, y = sr_heat)) + 
  geom_point() + 
  labs(title = paste0("r = ",round(cor(sharkdata$temperature, sharkdata$sr_heat),2)), 
       x = "actual temperature",y="self report measure of heat")
```

`r qbegin()`
Is $X$ (the number of ice-cream sales) a useful predictor of $Y$ (numbers of shark attacks) over and above $Z$ (temperature - measured on our self-reported heat scale)?  
<br>
```{r}
shark_mdl2a <- lm(shark_attacks ~ sr_heat + ice_cream_sales, data = sharkdata)
anova(shark_mdl2a)
```
<br>
What do you conclude?
`r qend()`

Moral of the story: be considerate of what exactly it is that you are measuring.  

<br>
This example was adapted from [Westfall and Yarkoni, 2020](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0152719) which provides a much more extensive discussion of incremental validity and type 1 error rates. 


<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
