---
title: "USMR 2021 Coursework"
author: "`r params$examnumber`"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
params:
  examnumber: CHANGE_TO_YOUR_EXAM_NUMBER
---

<!-- Beneath here is some code which will set everything up for you.  Please note that it will _empty your environment_ (removing any variables).   Anything that is needed to run your code should be explicitly set up below -->


```{r setup, include=FALSE}
# this line will mean that when you compile/knit your document,
# the code will not show, but the output (e.g., plots) will!
knitr::opts_chunk$set(echo = FALSE)
# this line will read in the data for you 
source("https://edin.ac/32Q9llb") 

# load any other packages that you require here:
library(tidyverse)

```



## Question 0

<!-- If you have run the R code above this point (you can do this now by pressing Ctrl-Alt-P, or running the chunk above) then your data will be in a dataframe called `dfsleep`. -->

Assess the data for any unusual values (given the context of the data description). Report any actions and decisions you plan to take regarding any such values, and use R to provide some descriptive statistics for the dataset.  


```{r cleaning, include = FALSE}
# Neither output nor code from this chunk will be shown in the compiled document. 


```


```{r descriptives}
# Code will not be shown from this chunk (because we set echo = FALSE in the very first chunk)
# the output from this code will be shown. 


```


## Question 1
Is self-reported sleep quality independent of academic role?  Perform an appropriate statistical test, and report and interpret the results in text.
  
```{r q1}


```




## Question 2 
Do smart-watch sleep quality scores relate to individuals' self-reports of perceived sleep quality? Does this relationship differ depending on academic role?


```{r q2}



```







## Question 3
Is sleep quality (as measured by smart-watch) influenced by the duration of screen usage in the hour prior to sleep?  Perform an analysis and report the results below.


```{r q3}


```







## Question 4
Is sleep quality (as measured by smart-watch) influenced by the duration of screen usage in the hour prior to sleep **after accounting for variation explained by sleep duration, age, and exercise**?  


```{r q4}


```





## Question 5 
Is the amount to which screen usage influences sleep quality different depending on whether screens are used in a light vs. a dark environment?  

```{r q5}


```





## Question 6
What factors increase the likelihood of reporting above-average sleep quality?  Here you should perform a more exploratory analysis using self-reported sleep quality as the outcome (dependent) variable. 

```{r q6}


```





### Extra 
Can you plot the predicted probabilities of an individual reporting above average sleep quality for 20 year old undergraduates who range from 0 to 12 hours of sleep?  


```{r q6extra}


```












