---
title: "USMR live R: week 11"
author: "USMR team"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```


We will first load the required package: tidyverse.

```{r}
library(tidyverse)
```

We will work with the class survey data:

```{r}
usmrsurv2 <- read_csv("https://uoepsy.github.io/data/usmrsurvey2.csv")
```

Check which variables were collected:

```{r}
names(usmrsurv2)
```

Give shorter names to some variables for ease of handling:

```{r}
names(usmrsurv2)[9:14] <- c("E","A","C","ES","I","LOC")

summary(usmrsurv2)
```

```{r}
summary(as.factor(usmrsurv2$gender))
```

Remember that the trick `1 * (condition`) transforms `TRUE/FALSE` to `1/0`

```{r}
usmrsurv2 <- usmrsurv2 %>% 
  filter(gender != "non-binary") %>% 
  mutate(is.female = 1*(gender=="female"))
```

Plot the indicator `is.female` as a function of `height`:

```{r}
usmrsurv2 %>% 
  ggplot(aes(x=height, y=is.female)) +
  geom_jitter(width=0, height=.1) +
  geom_smooth(method="glm", method.args=list(family=binomial), colour="black") +
  theme_classic()
```


## Model with height variable:

```{r}
model <- glm(is.female ~ height, family=binomial, data=usmrsurv2)

summary(model)
```

Intercept = log odds of being female when height = 0.

The interpretation of the intercept, involving height = 0, is not very meaningful. We can do better by mean-centering the height variable, leading to more meaningful estimates.


## Model with mean-centred height variable:

New plot:

```{r}
usmrsurv2 %>% 
  ggplot(aes(x=height-mean(height), y=is.female)) +
  geom_jitter(width=0, height=.1) +
  geom_smooth(method="glm", method.args=list(family=binomial), colour="black") +
  theme_classic()
```

First, create a new column in the data representing the mean-centred height:

```{r}
usmrsurv2 <- usmrsurv2 %>%
  mutate(height_mc = height - mean(height))
```

```{r}
model_mc <- glm(is.female ~ height_mc, family=binomial, data=usmrsurv2)

summary(model_mc)
```

logit estimate for a participant having a height of 10 above the mean height:

```{r}
3.3746 + (-0.4191) * 10
```

predicted logit = 3.3746 + (-0.4191) * 10 = -0.8164

```{r}
exp(-0.8164) / (1 + exp(-0.8164))
```

predicted prob = exp(logit) / (1 + exp(logit)) = exp(-0.8164) / (1 + exp(-0.8164)) = 0.3065

We can write a little function that takes any input and calls it x, and returns the transformation of x given by exp(x) / (1 + exp(x)). This is the transformation that goes from logit to prob, so we can call the function l2p:

```{r}
l2p <- function(x) {
  exp(x) / (1 + exp(x))
}
```

Apply this function to the logit computation:

```{r}
l2p(3.3746 + (-0.4191) * 10)
```

predicted prob = 0.3065


Let's verify the computation using the predict() function. It takes two arguments: (1) the fitted model, and (2) the new data at which you wish to obtain a prediction.

Create the new data, which contain the mean-centred height at which we want the prediction: 10. This is because we are interested in those having a height of 10 cm above the mean.

```{r}
newData <- tribble(
  ~height_mc,
  10
)
```

Predict the logit:

```{r}
predict(model_mc, newdata = newData)   # by default type = "link"
predict(model_mc, newdata = newData, type = "link")   # prediction in logit scale
l2p(-0.8164805)   # Transform logit to prob
```

Directly predict the prob:

```{r}
predict(model_mc, newdata = newData, type = "response")   # prediction in probability scale
```


