---
title: "Messy Data"
bibliography: references.bib
biblio-style: apalike
link-citations: yes
params: 
    SHOW_SOLS: FALSE
    TOGGLE: TRUE
---
    
```{r setup, include=FALSE}
source('assets/setup.R')
library(tidyverse)
library(patchwork)
```

:::red
**Preliminaries**  

1. Open Rstudio, make sure you have the USMR project open, and create a new RMarkdown document (giving it a title for this week).  
2. Download the data from [here]() (right click the link and choose "save target as"/"save file as" etc.)

:::

This week we're going to take a little detour away from statistics to get some practice wrangling and cleaning data in R.   

:::blue
**The 80/20 Rule**   
In many projects (both in acadmemic research & in other sectors), 80 percent of the time will be spent cleaning and organising data, and only 20% will be spent actually conducting statistical analyses^[A well designed study with a well-thought through data collection process can go a long way to remedy this!].
:::



```{r}
clickdat<-readRDS("~/Documents/git_archive/gcd_analysis/processed_data/gcd2_offline.RDS")

#clicks 
clickdat %>% select(subject, trial_no, obj_clicked) %>% 
    mutate(obj_clicked = fct_recode(obj_clicked, "Target"="Referent"),
           subject = paste0("subject_",subject)) %>%
    pivot_wider(names_from=trial_no, values_from=obj_clicked, names_prefix="Trial ") %>%
    #write.csv("data/trialcclicks.csv") %>%
    print

# conditions
clickdat %>% 
    mutate(cond = ifelse(condition=="No Cue", "_fluent.wav","_disfluent.wav"),
           audio = paste0("/files/",ref,cond)) %>%
    select(subject, trial_no, audio, rt) -> t

t$audio[grepl("umbrella_disfluent",t$audio)] <- "/files/umbrella_dislfuent.wav"
t$rt[sample(1:length(t$rt), size=4)]<-round(rnorm(4,13300,4050))
t$rt[sample(1:length(t$rt), size=4)]<-NA

#write.csv(t,"data/audio.csv",row.names=F)

head(clickdat)
summary(clickdat)
```


# Reading in data

## Different formats

TODO: table of functions for reading/writing data types.

:::frame

|  filetype|  description|  reading| writing|
|--:|--:|--:|
|  .csv|  2|  `read_csv()`<br>`read.csv()`<br>`read.table(..., sep = ",")`| `write.csv()`, |
|  .tsv|  3|  `read_tsv()`<br>`read.table(..., sep = "\t")`|
|  .txt|  3|  `read.table(..., sep = ...)`|

R files
|  filetype|  description|  reading| writing|
|  .RDS|  3|  `readRDS()`|
|  .RData|  3|  4|

Excel files
readxl


Other software
haven 
for instance with `read_spss()`/`read_sav()` and `read_sas()` for files from SPSS and SAS. 

Google sheets
googlesheets4 package
read_sheet()
requires authentication

:::

## Column names

```{r}
library(readxl)
read_xlsx("data/audio.xlsx")
audiodata <- read_xlsx("data/audio.xlsx")
audiodata
read_csv("data/trialcclicks.csv")
clicksdata <- read_csv("data/trialcclicks.csv", col_names = FALSE)
clicksdata <- clicksdata[,-1]
# clicksdata <- clicksdata %>% select(-1)

```

## Reshaping data 

```{r}
head(clicksdata)
clicksdata %>% 
    pivot_longer(cols = X3:X22, names_to = "trial_number", values_to = "object_clicked")
```




# Dealing with character strings

separate()
gsub()
substr()
    
# joining/merging

left_join/right_join
full_join
anti_join
semi_join
            
??? 

# identifying outliers
## dealing with outliers

# dealing with missingness

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
