---
title: "8A: Centering and Scaling"
params: 
    SHOW_SOLS: TRUE
    TOGGLE: TRUE
editor_options: 
  chunk_output_type: console
---



# centering/scaling/standardising

There are many transformations we can do to a continuous variable, but far and away the most common ones are _centering_ and _scaling._ 

:::sticky
__Centering__  

Centering simply means moving the entire distribution to be centered on some new value. We achieve this by subtracting our desired center from each value of a variable. 
A common option is to _mean center_ (i.e. to subtract the mean from each value).  

```{r}
dd = tibble(heights = rnorm(100,168,12),heightsc=heights-mean(heights))
ggplot(dd,aes(x=heights))+
  geom_histogram() +
ggplot(dd,aes(x=heightsc))+
  geom_histogram()
```


This makes our new values all relative to the mean. In the table below, the average age of our 4 people is 20.5.  
When we subtract 20.5 from each persons age, we get the mean-centered age (the `ageC` column below).  
```{r}
#| echo: false
tibble(
  name = randomNames::randomNames(4,which.names="first"), 
  age = c(18,20,22,22),
  ageC = age-mean(age)
)
```

We can center a variable on other things, such as the minimum or maximum value of the scale we are using, or some judiciously chosen value of interest.  

:::

:::sticky


:::

sticky scaling/standardising
subtract the mean, divide by the standard deviation
what does it do - makes sd 1, mean 0

consider Martin. he goes on a lot about people's heights. he is 182cm tall. and he has size 43 feet (in EU sizes)
is his tallness more unusual than his feet?  


# centering predictors  

changes intercept
why? changes what "0" represents in our data  

# scaling predictors

changes intercept and changes slope
why?
changes what "0" represents
changes what "moving 1" represents (now represents moving 1 SD)

# standardised coefficients  

puts everythng on the same magnitude 
makes qualitative comparisons more easy  


notation


# packages

BE CAREFUL THOUGH

