--- 
title: <svg width="50" height="50">
    <rect x="0" y="0" width="50" height="50"
    style="fill:#FCBB06;stroke:none;opacity:1" />
  </svg> 
  <br>
  Univariate Statistics and Methodology using R
author: "Department of Psychology, The University of Edinburgh"
date: "Academic year 2020-2021"
site: bookdown::bookdown_site
documentclass: book
bibliography: [references.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This book contains the lab materials for USMR at UoE."
---


```{r}
source("dropdowns.R")	
```


# Prerequisites

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
