--- 
title: <svg width="50" height="50">
    <rect x="0" y="0" width="50" height="50"
    style="fill:#FCBB06;stroke:none;opacity:1" />
  </svg> 
  <br>
  Univariate Statistics and Methodology using R
author: "Department of Psychology, The University of Edinburgh"
date: "Academic year 2020-2021"
site: bookdown::bookdown_site
documentclass: book
bibliography: [references.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This book contains the lab materials for USMR at UoE."
---

```{r setup, include=FALSE}
options(
  htmltools.dir.version = FALSE, formatR.indent = 2,
  width = 55, digits = 4, warnPartialMatchAttr = FALSE, warnPartialMatchDollar = FALSE
)
source("dropdowns.R")

knitr::opts_chunk$set(echo = T, message=F, warning=F, comment=NA, autodep=F, 
                      eval=T, cache.rebuild=F, cache=T, R.options=list(width=120), 
                      out.width = '40%', fig.asp = .5,
                      fig.width=8, fig.align = 'center')

```


# Course overview {-#chap-overview}

## Course description {-}

Univariate Statistics and Methodology in R (USMR) is a semester long crash-course aimed at providing Masters students in psychology with a competence in standard univariate methodology and analysis using R. Design and analysis are taught under a unifying framework which shows a) how research problems and design should inform which statistical method to use and b) that many statistical methods are special cases of a more general model. This course will introduce you to statistical modelling and empower you with tools to analyse richer data and answer a broader set of research questions of interest to you.

This course introduces you to statistics and R software. We begin with introductions to the use of statistical methods in research and to R for statistics. We then move to learning about inferential statistics, covering concepts such as hypothesis testing, Type I vs. Type II errors, p-values and power. Common tests such as t-tests and chi-squared tests are introduced before building to a thorough explanation of simple linear regression and its application and assumptions. Extending this to multiple regression with interaction terms and categorical coding schemes, the course culminates in introducing the generalised linear model and the different families of models.

## Team {-}

- Professor Martin Corley: martin.corley@ed.ac.uk
- Dr Josiah King: ug.ppls.stats@ed.ac.uk
- Dr Umberto Noe: ug.ppls.stats@ed.ac.uk
- And not forgetting your friendly tutoring team! Ask them anything! 

## Schedule {-}

| Week| Lecture | Lab |
|:------|:------|:------|
| 1     | Introduction to statistics and research methods | Introduction to R<br>Collecting data<br>Types of data |
| 2   | Logic, sets, probability, sampling, descriptive statistics, and variables  | Visualising and describing distributions<br> Visualising and describing relationships  |
| 3   | Functions, normal distribution, central limit theorem, data visualisation  | Probability theory<br>Random variables<br>Sampling variability and sampling distributions<br>Bias-variance trade-off |
| 4   | Models, hypotheses, probability distributions, significance testing  | Bootstrap & Confidence Intervals<br>Hypothesis testing with the p-value approach<br>Hypothesis testing with the critical values approach<br>Hypothesis testing & Confidence Intervals<br>Making decisions - Effect sizes, Power, Errors |
| 5   | Statistical tests: $z$, $t$, $\chi^2$, $F$  |  Test for two mean (independent samples)<br>Test for one mean and test for two means (paired samples)<br>Chi-square test|
| Break | | |
| 6   | GLM 1: Correlation and bivariate regression  | Correlation (different types)<br>Move to regression<br>Simple linear regression<br>Interpreting coefficients<br>Assumptions and violations |
| 7   | GLM 2: Multiple regression and model checking  | Multiple linear regression<br>Interpreting coefficients<br>Assumptions<br>Model selection<br>Model issues |
| 8   | GLM 3: Interactions, effects coding, standardisation  |  Interactions (quant $\times$ quant)<br>Interactions (mixed)<br>ANOVA as special case<br>Coding (dummy, effect)<br>Interactions (cat $\times$ cat)|
| 9   | GLM 4: Generalising the GLM - link functions and estimation |  Logistic regression<br>Assumptions<br>Multinomial Logistic<br>Other GLM|
| 10  | GLM and ANOVA  | Recap - common tests as linear model |

  


## Textbook {-}
  
The course textbook is _Learning Statistics with R_ (version 0.4 or higher) by D Navarro.  
Download the book (version 0.6; [PDF](https://learningstatisticswithr.com/lsr-0.6.pdf)) or purchase a printed copy for around £20)  
All datasets are available at: https://learningstatisticswithr.com  



```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```