---
title: "Test of learnr + webassemby"
link-citations: true
params: 
    SHOW_SOLS: TRUE
    TOGGLE: TRUE
---

```{r setup, include=FALSE}
source('assets/setup.R')
library(xaringanExtra)
xaringanExtra::use_panelset()
library(webexercises)
```

{{< include ./_extensions/r-wasm/live/_knitr.qmd >}} 
{{< include ./_extensions/r-wasm/live/_gradethis.qmd >}}

## Introduction


we have content.    

read this. 
```{webr}
#| include: false
x = rnorm(100)
y = rnorm(100, 1*x, 1)
df1 <- data.frame(x,y)

webr::eval_js('
  async function submitForm(data) {
    const formData = new FormData();
    formData.append("entry.839205657", data.field1);
    
    try {
      const response = await fetch("https://docs.google.com/forms/d/e/1FAIpQLSekV6BBmR3EWhTERZOK-boMsW-lC1HeKN76gWG95q7FeTE3vA/formResponse", {
        method: "POST",
        body: formData,
        mode: "no-cors"
      });
      return "Form submitted successfully";
    } catch (error) {
      return "Error: " + error.message;
    }
  }
')
submit <- function(response=0){
  webr::eval_js(paste0('submitForm({
  field1: "', response, '"})'))
  print("Thank you for your feedback!")
}
```


above here is a webr chunk with include: false. Because the webr stuff is basically just evaluated in a single env for the page, it might be worth having an initial chunk to pre-populate the page with loading the relevant libraries & data etc.  


here is something you can play with:  
can you guess what the coefficient will be? 
```{webr}
#| caption: Simple regression
x <- rnorm(100)
y <- rnorm(100, 1*x, 1)
df <- data.frame(x,y)

lm(y ~ x, data = df) |>
  summary()
```

What about this one? where is df1 from? the author has made it for you already! 
```{webr}
#| caption: Simple regression
lm(y ~ x, data = df1) |> coef()
```



blah blah.   

here's a question.. what is 3+4?  
```{webr}
#| exercise: initq
#| setup: true

```

```{webr}
#| caption: Exercise
#| exercise: initq

```

::: {.hint exercise="initq"}
::: {.callout-note collapse="false"}
#### Hint 1

<!-- hint text here -->

```{r}
#| eval: false
3+?
```
:::
:::

::: {.solution exercise="initq"}
::: {.callout-tip collapse="false"}
#### Solution

<!-- Solution text here -->

```{webr}
#| exercise: initq
#| solution: true
3+4
```
:::
::::

```{webr}
#| exercise: initq
#| check: true
#| class: wait
gradethis::grade_this_code()
```


# part 2

hello you made it to part 2 because you correctly calculated 3+4.  

if we wanted to link people directly to pages that avoid this incremental reveal, then we just use:  

- *with* incremental reveal: https://uoepsy.github.io/lm/00activetest.html
- *without* incremental reveal: https://uoepsy.github.io/lm/00activetest.html?displayall
 
here are some more types of question:  

```{r}
#| echo: false
#| results: 'asis'
# library(webexercises)
qm1 <- c(
  3, 
  4,
  1,
  answer = 7
  )
```

| Question | &nbsp; |
| -------- | ------ |
| what was 3+4? | `{r} I(mcq(qm1))` |
| 3 + 4 = 5 | `{r} I(torf(FALSE))` |

- [ ] Continue to the next section


# part 3

and here we are.  


## feedback? 

How did you find this reading/resource?  

Use the following numbers to submit quick feedback (i.e., `submit(1)`, or leave a custom message for us `submit("Hello my friends!")`):  

1 - "I got to the end! Woohoo!"
2 - "This was boring. i knew it all already"
3 - "This helped. i think i understand things a little better now"
4 - "This just made things more confusing"

```{webr}
submit()
```


