---
title: "Violated Assumptions & Correlated Predictors"
bibliography: references.bib
biblio-style: apalike
link-citations: yes
params: 
    SHOW_SOLS: TRUE
    TOGGLE: FALSE
---

```{r setup, include=FALSE}
source('assets/setup.R')
library(tidyverse)
library(patchwork)
```


# Violated Assumptions

At the high-level view, we want to see no evidence of any systematic patterns in our residuals. We want them to look more or less like random noise. This means we want them to be normally distributed, with a mean of zero. Furthermore, we want the mean and the variance around that mean to remain constant across the values of the linear predictor (the fitted model).  

`r optbegin("What does it look like?",olabel=FALSE,toggle=params$TOGGLE)`
```{r echo=FALSE, fig.height=4}
library(ggdist)
library(tidyverse)
df<-tibble(x=runif(1000,1,10),xr = round(x), y=1*x+rnorm(1000))
df$y2 <- resid(lm(y~x,df))
#df$y[df$x==6]<-6+rnorm(100,0,3)
df %>% group_by(xr) %>% summarise(m=mean(y2), s=sd(y2)) -> dfd
p1 <- ggplot(df, aes(x=x,y=y))+
  geom_point()+
  geom_smooth(method="lm",se=F, fullrange=T)+
  scale_x_continuous("x",1:10, breaks=seq(1,10,2))+
  scale_y_continuous("y")
p2 <- ggplot(df, aes(x=xr,y=y2))+
  geom_jitter(height=0,width=1, alpha=.3)+
  stat_dist_halfeye(inherit.aes=F,data=dfd, aes(x=xr,dist="norm",arg1=m,arg2=s),alpha=.6, fill="orange")+
  #geom_smooth(method="lm",se=F, fullrange=T)+
  scale_x_continuous("x",1:10, breaks=seq(1,10,2))+
  scale_y_continuous("residuals")

p1 + p2
```
`r optend()`

`r optbegin("What does it <b>not</b> look like?",olabel=FALSE,toggle=params$TOGGLE)`
Things look a bit wrong (like there is something systematic that we haven't accounted for), when our residuals do not have mean zero:  
```{r echo=FALSE, fig.height=4}
library(ggdist)
library(tidyverse)
tibble(x = runif(1000,1,10),
       xr = round(x),
       s = abs(5-x)*2,
       e = map_dbl(s,~rnorm(1,0,1)),
       y = x + s + e) -> df
df$y2 <- resid(lm(y~x,df))
#df$y[df$x==6]<-6+rnorm(100,0,3)
df %>% group_by(xr) %>% summarise(m=mean(y2), s=sd(y2)) -> dfd
p1 <- ggplot(df, aes(x=x,y=y))+
  geom_point()+
  geom_smooth(method="lm",se=F, fullrange=T)+
  scale_x_continuous("x",1:10, breaks=seq(1,10,2))+
  scale_y_continuous("y")
p2 <- ggplot(df, aes(x=xr,y=y2))+
  geom_jitter(height=0,width=1, alpha=.3)+
  stat_dist_halfeye(inherit.aes=F,data=dfd, aes(x=xr,dist="norm",arg1=m,arg2=s),alpha=.6, fill="orange")+
  scale_x_continuous("x",1:10, breaks=seq(1,10,2))+
  scale_y_continuous("residuals")

p1 + p2
```

Or do not have constant variance: 
```{r echo=FALSE, fig.height=4}
library(ggdist)
library(tidyverse)
tibble(x = runif(1000,1,10),
       xr = round(x),
       s = abs(x)/2,
       e = map_dbl(s,~rnorm(1,0,.)),
       y = x + e) -> df
df$y2 <- resid(lm(y~x,df))
#df$y[df$x==6]<-6+rnorm(100,0,3)
df %>% group_by(xr) %>% summarise(m=mean(y2), s=sd(y2)) -> dfd
p1 <- ggplot(df, aes(x=x,y=y))+
  geom_point()+
  geom_smooth(method="lm",se=F, fullrange=T)+
  scale_x_continuous("x",1:10, breaks=seq(1,10,2))+
  scale_y_continuous("y")
p2 <- ggplot(df, aes(x=xr,y=y2))+
  geom_jitter(height=0,width=1, alpha=.3)+
  stat_dist_halfeye(inherit.aes=F,data=dfd, aes(x=xr,dist="norm",arg1=m,arg2=s),alpha=.6, fill="orange")+
  scale_x_continuous("x",1:10, breaks=seq(1,10,2))+
  scale_y_continuous("residuals")

p1 + p2
```
`r optend()`

But more often than not, when face with _real_ data (rather than the simulated datasets used in textbooks and intro stats courses!), things rarely look perfect.  

## Model Mis-specification

- might there be some non-linearity in the relationship you are studying? Non-linearity may also take the form of an interaction - the relationship $y \sim x$ is not linear if it _depends on_ the value of $z$. 

- might there be some ommitted variable 

- might 

## Transformations

## Bootstrapping

## PCA



<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
