---
title: "<b>Week 4: More Tests</b>"
subtitle: "Univariate Statistics and Methodology using R"
author: "Martin Corley"
institute: "Department of Psychology<br/>The University of Edinburgh"
date: "AY 2020-2021"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: 
      - xaringan-themer.css
      - mc_libs/tweaks.css
    nature:
      beforeInit: "mc_libs/macros.js"
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(digits=4,scipen=2)
options(knitr.table.format="html")
xaringanExtra::use_xaringan_extra(c("tile_view","animate_css","tachyons"))
xaringanExtra::use_extra_styles(
  mute_unhighlighted_code = FALSE
)
library(knitr)
library(tidyverse)
library(ggplot2)
source('R/pres_theme.R')
knitr::opts_chunk$set(
  dev = "svg",
  warning = FALSE,
  message = FALSE,
  cache = TRUE
)
source('R/myfuncs.R')
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_accent(
  #base_color = "#0F4C81", # DAPR1
  # base_color = "#BF1932", # DAPR2
  # base_color = "#88B04B", # DAPR3 
  base_color = "#FCBB06", # USMR
  # base_color = "#a41ae4", # MSMR
  header_color = "#000000",
  header_font_google = google_font("Source Sans Pro"),
  header_font_weight = 400,
  text_font_google = google_font("Source Sans Pro", "400", "400i", "600", "600i"),
  code_font_google = google_font("Source Code Pro")
)
```

class: inverse, center, middle

# Part 1

### R:  The Big Ideas

---
class: inverse, center, middle, animated, rollIn

# End of Part 1

---
class: inverse, center, middle

# Part 2

### Statistics:  The Big Ideas

---
class: inverse, center, middle, animated, rollIn

# End of Part 2

---
- the big ideas include the idea that if you know the chance distribution, you
  can work out probability
  
- in Part 3 maybe simulate a x-sq distribution??
  + live R?

- any other tests?
---
class: inverse, center, middle

# Part 3

### Statistics:  The $\chi^2$ tesst

---
- combine "big ideas" -- simulate, use plot() on categories -> mosaic
---
class: inverse, center, middle, animated, rollIn

# End

---
# Acknowledgements

- icons by Diego Lavecchia from the [Noun Project](https://thenounproject.com/)