---
title: "A Coursework-Like Question"
subtitle: Univariate Statistics and Methodology using R
output:
  tufte::tufte_html: default
  tufte::tufte_handout: default
---


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
require(tidyverse)

# knitr::opts_chunk$set(echo = FALSE, eval=FALSE, warning=FALSE, message=FALSE)
# # For solutions:
knitr::opts_chunk$set(echo = FALSE, eval=FALSE, warning=FALSE, message=FALSE)

load(url("https://uoepsy.github.io/data/usmr_1920_assignment.RData"))
```

# Assignment  

For this assignment, you will be expected to clean and analyse a data set, and report your findings.

## Motoring offences dataset  

Two datasets can be loaded from the following url:  
```{r, eval=FALSE,echo=TRUE}
load(url("https://uoepsy.github.io/data/usmr_1920_assignment.RData"))
```

The data provided contains information about the nature and circumstances of motorists stopped and breathalysed by the Police.  
Data is collected every time that driver is stopped by the Police and breathalysed. Records indicate the speed at which the driver is travelling when they are stopped, and the blood alcohol content of the driver when measured via breathalyser. Information is also captured on the age and prior motoring offences of the driver, and whether the incident occurred at day or night. Police officers may have had reasons for stopping drivers other than presuming them to be intoxicated (for instance, someone who is stopped for speeding may subsequently be breathalysed if they are deemed to be acting unusually).  

Each time a police officer stops a motorist, an incident ID is created. A separate database used primarily for administrative purposes includes records of which officer (recorded as initials) attends which incidents.  

```{r, echo=FALSE, results="asis", eval=TRUE}
dict<-matrix(c(names(drinkdriving),"officer",
         "Age of driver (in years)","Whether or not the incident occurred at night",
         "Offence code for any prior motoring offences","Speed when stopped by police (mph)",
         "Blood Alcohol Content (%) as measured by breathalyser","Outcome of stop ('fine','warning')",
         "ID of incident","Officer attending (initials)"), nrow=(length(names(drinkdriving))+1))
colnames(dict)<-c("Variable","Description")
off<-matrix(c("N","DR50","DR80","CD..","PL..","SP..","TS..",
         "No prior offence","In charge of veh. while unfit through drink",
         "In charge of veh. while unfit through drugs","Careless Driving offences ...",
         "Driving without 'L' Plates","Speeding offences ...",
         "Traffic direction and signs offences ..."), nrow=7)
colnames(off)<-c("Offence code","Description")

cat("## Data dictionary  \n")
knitr::kable(dict)
# cat("## Offence Codes  \n")
# knitr::kable(off)
```

# Sample Question:  Driving speeds, night vs. day  

Does time of day and speed of driving predict the blood alcohol content over and above driver's age? Fit appropriate model(s) to test this question, and report the results (you may add a figure or table if appropriate).

