---
title: "<b>Week 1: Introduction</b>"
subtitle: "Univariate Statistics and Methodology using R"
author: "Martin Corley"
institute: "Department of Psychology<br/>The University of Edinburgh"
date: "AY 2020-2021"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: 
      - xaringan-themer.css
      - mc_libs/tweaks.css
    nature:
      beforeInit: "mc_libs/macros.js"
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(digits=4,scipen=2)
xaringanExtra::use_xaringan_extra(c("tile_view","animate_css","tachyons"))
xaringanExtra::use_extra_styles(
  mute_unhighlighted_code = FALSE
)
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_accent(
  #base_color = "#0F4C81", # DAPR1
  # base_color = "#BF1932", # DAPR2
  # base_color = "#88B04B", # DAPR3 
  base_color = "#FCBB06", # USMR
  # base_color = "#a41ae4", # MSMR
  header_color = "#000000",
  header_font_google = google_font("Source Sans Pro"),
  header_font_weight = 400,
  text_font_google = google_font("Source Sans Pro", "400", "400i", "600", "600i"),
  code_font_google = google_font("Source Code Pro")
)
```


# Shape of the Course


.flex.items-center[.w-20.pa2[![:scale 40%](img/lecture.svg)] .f2.w-80.pa2[Lectures (online)]]

.flex.items-center[.w-20.pa2[![:scale 40%](img/labs.svg)] .f2.w-80.pa2[Labs (instructions online)]]

.flex.items-center[.w-20.pa2[![:scale 40%](img/forum.svg)] .f2.w-80.pa2[Discussion and help]]

???
As soon as we can, we'll offer discussions and help on campus.

We might be able to offer labs on campus as the situation changes, and we hope to do so.

Lectures are likely to remain online throughout the course.

---
# Lectures and Labs

<!-- FIXME -->


.pull-left[
### Lectures released on .red[Mondays at 11am]
]
.pull-right[
### Labs released on .red[Tuesdays at 11am]
]
--
.pt3[
.pull-left[
- lectures include videos, activities
- no need to use RStudio:  everything you need online
]
.pull-right[
- labs make use of **RStudio** (an environment for **R**)
- can either install on your own laptop, or use our **server**
]]

--

<br/>

.pt3[
### The best way to do well on this course
- keep on top of the activities
- let us know _early_ (discussion fora, email) if you're struggling

]


---
# We are here to help you

.flex.items-center[
<!-- PLAYMOPIC --> <!-- Use the forums; come and meet us there -->
]

---
# Acknowledgements

- icons by Becris, tezar tantula, ProSymbols from the [Noun Project](https://thenounproject.com/)